---
title: "Statistical Simulation and Inferential Data Analysis"
subtitle: "Inferential Statistics: Course Project"
author: "John Snyder"
date: "August 26, 2018"
output: 
    pdf_document:
      number_sections: yes
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```

# Overview

The goal of this document is twofold. In the first section, simulation is used to demonstrate the Central Limit Theorem.  In the second section, basic statistical inference techniques are used to analyse the `ToothGrowth` data set.

# Central Limit Theorem

The central Limit Theorem states

## Simulations

```{r,plotExpSamples, fig.height=3, fig.width=4, fig.cap="Exponential Distribution Simulated Sample and PDF", }
library(ggplot2)

lambda <- 0.2
n <- 40
m <- 1000

set.seed(2018) # for reproducibility
expSamples <- rexp(n*m, lambda)
expMatrix <- matrix(expSamples, ncol=n, nrow=m)


ggplot(data.frame(exp=expSamples), aes(x=exp)) +
    geom_histogram(color="black", bins=50, aes(y=..density..)) +
    stat_function(fun=dexp,
                  color="red",
                  size=1,
                  args=list(rate=lambda))+
    labs(y="Probability Density",
         x="Observed Value")
```

## Sample vs. Theoretical Mean

Code for the graphs below is based on the code provided in [a StackExchange question](https://stackoverflow.com/questions/29182228/plotting-normal-curve-over-histogram-using-ggplot2-code-produces-straight-line)

```{r mean, fig.height=3, fig.width=4, fig.cap="Sample of 1000 means from the exponential distribution, n=40 for each sample"}
# Calculate the means of the 1000 Samples
expMeans <- apply(expMatrix, 1, mean)

# Plot the distribution of means, along with the theoretical mean for
# the exponential distribution, the measured mean, and the PDF for 
# a normal distribution with a mean equal to the theoretical mean and
# a standard deviation equal to our sample standard deviation.
ggplot(data.frame(means = expMeans), aes(x=means)) +
    geom_histogram(color = "black", bins=50, aes(y=..density..)) +
    stat_function(fun=dnorm, 
                  color="red",
                  size=2,
                  args = list(mean=1/lambda, 
                              sd=sd(expMeans))) +
    geom_vline(xintercept = 1/lambda, color="red", size=2)+
    geom_vline(xintercept = mean(expMeans), color="blue", size=2) +
    labs(y="Probability Density",
         x="Sample Mean (n=40)")
```

## Sample vs. Theoretical Variance

```{r sd, fig.height=3, fig.width=4, fig.cap="Sample of 1000 standard deviations from the exponential distribution, n=40 for each sample"}
expSDs <- apply(expMatrix, 1, sd)

ggplot(data.frame(sd = expSDs), aes(x=sd)) +
    geom_histogram(color = "black", bins=50, aes(y=..density..)) +
    stat_function(fun=dnorm, 
                  color="red",
                  size=2,
                  args = list(mean=1/lambda, 
                              sd=sd(expSDs))) +
    geom_vline(xintercept = 1/lambda, color="red", size=2)+
    geom_vline(xintercept = mean(expSDs), color="blue", size=2) +
    labs(y="Probability Density",
         x="Sample Standard Deviation (n=40)")
```

# Inferential Data Analysis

```{r}
library(dplyr)

data("ToothGrowth")
summary <- ToothGrowth %>% 
           group_by(dose, supp) %>% 
           summarize(mean=mean(len), sd=sd(len), n=n())

knitr::kable(summary, caption = "Summary of Tooth Growth Data")
```

```{r}
ggplot(ToothGrowth, aes(x=factor(dose), y=len, fill=supp)) +
  geom_boxplot()
```

```{r}
signif <- ToothGrowth %>% 
          group_by(dose) %>% 
          summarize(p = t.test(
            len[supp=="OJ"],
            len[supp=="VC"],
            alternative = "two.sided",
            paired = FALSE)$p.value)
knitr::kable(signif, caption="Significance Level by Dose")
```          
         